Preventing a Nuclear ISIL: \\Recommendations for United States Policy

### 2015-05-11 draft.tex
Added in Conclusions section.

\usepackage{pdflscape}

@@

Eric Harvey, your assigned section is still missing after being 3 weeks late! Get it to us yesterday!

@@

\chapter{???????  Do final consistency checks for ISIS/ISIL, US/U.S., apostrophes, SLD/SLoD, quote marks  ???????????}

@@

Waiting on Denia?

@@

@@

\begin{table}[h]

@@

>>>>

\usepackage{pdflscape}
\usepackage{afterpage}

@@

Placeholder for Eric Harvey's missing section on orphan / medical sources policy.

@@

% \chapter{???????  Do final consistency checks for ISIS/ISIL, US/U.S., apostrophes, SLD/SLoD, quote marks  ???????????}

@@

Though the probability of ISIL acquiring nuclear weapons, a radiological dispersal device, or special nuclear material may be small, it is not zero. If ISIL indeed acquired any of these through the possible pathways outlined in this report (or otherwise), the consequences would be dire.

This report has illustrated the details of policy recommendations focused on preventing the low-probability / high-consequence event of ISIL becoming a nuclear threat. To target all possible pathways, this report has made policy recommendations - new ones, as well as additions to existing policy - in an effort to prevent weapons theft or purchase, the strengthening of ISIL foundational capabilities for indigenous development, or the acquisition of radiological material for use in a radiological dispersal device. These policies are  summarized here:

\begin{easylist}[enumerate]
& Policies specifically targeting \textbf{indigenous development:}
&& Targeting of foundational capabilities	
&&& Income 
&&& People and expertise 
&&& Infrastructure
&& IAEA monitoring 
&&& Additional Protocol – push for more signatures 
&&& International and Trafficking Database – push for increased cooperation
& Policies specifically targeting \textbf{proliferation through theft/purchase:}
&& IAEA monitoring 
&&& International and Trafficking Database – push for increased cooperation
&&& Additional Protocol – push for more signatures 
&& Global Threat Reduction Initiative (GTRI) 
&&& Push for the cooperation of more nations
&& Strengthening of multilateral security agreements
&& Nuclear Forensics Attribution Act (NFAA)
&&& Explicit declaratory policies, explicit deterrence 
&& United Nations Security Council Resolution (UNSCR) 1540 
&&& Additional provisions and security expert group
&& Second Line of Defense (SLD)
&&& Additional foreign ports
& Policies specifically targeting \textbf{delivery/staging:}
&& Military Force Protection Condition (FPCON)
&&& Increased military base safety
&& Advanced Radiation Portal Monitor (RPM) Plan
&&& Increased domestic port monitoring and advanced monitor development
&& Second Line of Defense (SLD)
&&& Monitor additional foreign ports
\end{easylist}

As with the development of any policy, decisions are based on tradeoffs. It is important to understand the nature of diminishing returns when targeting efforts and funds to prevent an ISIL nuclear threat. This is true not only for crafting US policy, but also for understanding the intentions and motivations of ISIL and assessing the probability of their efforts to acquire nuclear material or devices. It is imperative to match U.S. policy with an understanding of the likelihood that ISIL will pursue one or more of these possible nuclear routes.

Furthermore, the caveats in implementing these policies must be understood. The realization of some of the policy recommendations might be diplomatically difficult in light of current political relationships, keeping in mind nuclear cooperation agreements (or lack thereof) with other countries, and the relationships between states relevant to this conflict. It is important to continue to encourage the recognition of the diplomatic difficulty in the policymaking and recommendation process.

The recommendations that have been illustrated in this report rest on the assumption that a nuclear ISIL needs to be prevented at any cost. A dissenting viewpoint has been offered and explored, and recommends focusing efforts on dismantling ISIL as a movement instead of dedicating disproportionate resources to an unknown probability of nuclear acquisition. Regardless of the philosophy on how to allocate funding to implement policy options, it is at the very least prudent to spend efforts on reducing knowledge uncertainty of possible ISIL-related nuclear event. Part of the low-hanging fruit of increasing preparedness and U.S. response capability (especially keeping in mind the cost, feasibility, and tradeoffs of policy implementation) is the assessment of possible nuclear outcomes, such as this report has done. However large or small the perceived risk, the work contained in this report contributes to the knowledge of prevention and response to the possibility of a nuclear ISIL.

Although the discussion about the allocation of efforts continues among policymakers, the Obama administration's National Security Strategy in 2010 has strongly stated its position: \enquote{This Administration has no greater responsibility than the safety and security of the American people. And there is no greater threat to the American people than weapons of mass destruction, particularly the danger posed by the pursuit of nuclear weapons by violent extremists and their proliferation to additional states} \cite{Obama2010}. The work in this report supports the national security goals of the current administration if the recommended strategies are implemented.

% \section{??????? ORDERED LIST OF RECOMMENDATIONS HERE ????????}

@@

% \afterpage{

@@

\begin{table}

@@

% }
% \nopagebreak[4]
